---
title: "fk_infographic_data"
author: "Megan Hepner"
date: "August 19, 2017"
output: html_document
---

```{r librarys}
#rm(list = ls())
library(knitr)
library(tidyverse)
library(rvc)
library(vegan)
library(dygraphs)
library(grDevices)
library(RColorBrewer)

```

```{r read domain rvc data}

RVCdata_FK = read_rds("C:/Users/mhepn/Documents/RVC/big_csv/RVCdata_FK.rds") #473 MB

domain_abun = read_csv("C:/Users/mhepn/Documents/RVC/big_csv/abundance_domain/domain_fk_abun.csv") #1.3 MB

domain_biomass = read_csv("prep/data/domain_fk_biomass.csv")

domain_density = read_csv("prep/data/domain_fk_density.csv")

```

```{r forage fish}

forage_fk_abun_csv     = "prep/data/forage_fk_abun.csv"        #abundance
forage_fk_den_csv      = "prep/data/forage_fk_den.csv"           #density
forage_fk_bio_csv      = "prep/data/forage_fk_bio_merged.csv"    #biomass 
forage_fk_diversity_csv= "prep/data/forage_fk_diversity.csv"   #diversity 

if (!all(file.exists(
    forage_fk_abun_csv, forage_fk_den_csv, forage_fk_bio_csv,forage_fk_diversity_csv))){
  
  forage_fish_spp_list= c("HYP HARR", "ATH STIP", "JEN SPE.", "HAR JAGU", "ANC LYOL", "INE VITT", "HEM BRAS", "SAR AURI", "HAR HUME", "CHR ATHE")

  #abundance
  forage_fk_abun = getDomainAbundance(RVCdata_FK, forage_fish_spp_list)
  write_csv(forage_fk_abun, "prep/data/rvc_trophic_groups/forage_fk_abun.csv")
 
  #density
  forage_fk_den = getDomainDensity(RVCdata_FK, forage_fish_spp_list)
  write_csv(forage_fk_den, "prep/data/rvc_trophic_groups/forage_fk_den.csv")
  
  #biomass
  forage_fk_bio = getDomainBiomass(RVCdata_FK, forage_fish_spp_list, RVCdata_FK$taxonomic_data)
  write_csv(forage_fk_bio, "prep/data/rvc_trophic_groups/forage_fk_bio_merged.csv")
 
  #fk diversity metrics 
  forage_fk_abundance = forage_fk_abun %>% 
    select(YEAR, protected_status, SPECIES_CD, abundance) %>%  
    group_by(YEAR, protected_status) %>% #48 groups 
    nest(-YEAR) %>%  
    mutate(
      data_wide = map(data, ~ spread(data=.x, SPECIES_CD, abundance, fill = 0)), 
      #species richness 
      richness = map(
        data_wide, 
        function(x) specnumber(x)),
      #simpson diversity as effective number of species 
      simpson = map( 
        data_wide,
        function(x) 1/(1 - diversity(x, index = 'simpson'))),
      #shannon diversity as effective number of species 
      shannon = map(
        data_wide,
        function(x) exp(diversity(x, index = 'shannon')))) %>%
    unnest(richness, simpson, shannon)
  
  forage_fk_diversity = forage_fk_abundance %>%
    select(YEAR, protected_status, richness, shannon, simpson) 
  write_csv(forage_fk_diversity, "big_csv/trophic_groups/diversity/forage_fk_diversity.csv")
} else{
  forage_fk_abun = read_csv("prep/data/rvc_trophic_groups/forage_fk_abun.csv")
  forage_fk_den = read_csv("prep/data/rvc_trophic_groups/forage_fk_den.csv")
  forage_den = forage_fk_den %>%
    mutate(var_den = var) %>%
    select(density, var_den)
  
  forage_fk_bio = read_csv("prep/data/rvc_trophic_groups/forage_fk_bio_merged.csv")
}

