---
title: "fk_infographic_data"
author: "Megan Hepner"
date: "August 19, 2017"
output: html_document
---

```{r librarys}
#rm(list = ls())
library(knitr)
library(tidyverse)
library(rvc)
library(vegan)
library(dygraphs)
library(grDevices)
library(RColorBrewer)

```

```{r read domain rvc data}

if(!file.exists("C:/Users/mhepn/Documents/RVC/big_csv/RVCdata_FK.rds")){
  RVCdata_FK <- getRvcData(1999:2016, "FLA KEYS", server = 'https://www.sefsc.noaa.gov/rvc_analysis20/')
  write_rds(RVCdata_FK, 'big_csv/RVCdata_FK.rds')
  
}else{
  RVCdata_FK = read_rds("C:/Users/mhepn/Documents/RVC/big_csv/RVCdata_FK.rds") #473 MB
}

```

### Forage fish

| Scientific name               | Common name         | RVC code  |  
|:------------------------------|:--------------------|:----------|
|*Hypoatherina harringtonensis* | reef silverside	    | HYP HARR  |
|*Atherinomorus stipes*	        | hardhead silverside	| ATH STIP  |
|*Jenkinsia sp.*	              | herring species	    | JEN SPE.  |
|*Harengula jaguana*	          | scaled sardine	    | HAR JAGU  |
|*Anchoa lyolepis*	            | dusky anchovy	      | ANC LYOL  |
|*Inermia vittata*	            | boga	              | INE VITT  |
|*Hemiramphus brasiliensis*	    | ballyhoo	          | HEM BRAS  |
|*Sardinella aurita*	          | spanish sardine	    | SAR AURI  |
|*Harengula humeralis*	        | redear sardine	    | HAR HUME  |
|*Chriodorus atherinoides*	    | hardhead halfbeak	  | CHR ATHE  |

```{r forage fish}

if(!all(file.exists("prep/data/rvc_trophic_groups/forage_abun.csv",
                    "prep/data/rvc_trophic_groups/forage_den.csv",
                    "prep/data/rvc_trophic_groups/forage_bio.csv",
                    "prep/data/rvc_trophic_groups/forage_div.csv"))){
  
  forage_fk_abun_csv     = "prep/data/forage_fk_abun.csv"        #abundance
  forage_fk_den_csv      = "prep/data/forage_fk_den.csv"           #density
  forage_fk_bio_csv      = "prep/data/forage_fk_bio_merged.csv"    #biomass 
  forage_fk_diversity_csv= "prep/data/forage_fk_diversity.csv"   #diversity 
  
  if (!all(file.exists(
    forage_fk_abun_csv, forage_fk_den_csv, forage_fk_bio_csv,forage_fk_diversity_csv))){
    
    forage_fish_spp_list= c("HYP HARR", "ATH STIP", "JEN SPE.", "HAR JAGU", "ANC LYOL", "INE VITT", "HEM BRAS", "SAR AURI", "HAR HUME", "CHR ATHE")
    
    #abundance
    forage_fk_abun = getDomainAbundance(RVCdata_FK, forage_fish_spp_list)
    write_csv(forage_fk_abun, "prep/data/rvc_trophic_groups/forage_fk_abun.csv")
    
    #density
    forage_fk_den = getDomainDensity(RVCdata_FK, forage_fish_spp_list)
    write_csv(forage_fk_den, "prep/data/rvc_trophic_groups/forage_fk_den.csv")
    
    #biomass
    forage_fk_bio = getDomainBiomass(RVCdata_FK, forage_fish_spp_list, RVCdata_FK$taxonomic_data)
    write_csv(forage_fk_bio, "prep/data/rvc_trophic_groups/forage_fk_bio_merged.csv")
    
    #fk diversity metrics 
    forage_fk_abundance = forage_fk_abun %>% 
      select(YEAR, protected_status, SPECIES_CD, abundance) %>%  
      group_by(YEAR, protected_status) %>% #48 groups 
      nest(-YEAR) %>%  
      mutate(
        data_wide = map(data, ~ spread(data=.x, SPECIES_CD, abundance, fill = 0)), 
        #species richness 
        richness = map(
          data_wide, 
          function(x) specnumber(x)),
        #simpson diversity as effective number of species 
        simpson = map( 
          data_wide,
          function(x) 1/(1 - diversity(x, index = 'simpson'))),
        #shannon diversity as effective number of species 
        shannon = map(
          data_wide,
          function(x) exp(diversity(x, index = 'shannon')))) %>%
      unnest(richness, simpson, shannon)
    
    forage_fk_diversity = forage_fk_abundance %>%
      select(YEAR, protected_status, richness, shannon, simpson) 
    write_csv(forage_fk_diversity, "prep/data/rvc_trophic_groups/forage_fk_diversity.csv")
    
  } else{
    
    forage_fk_abun = read_csv("prep/data/rvc_trophic_groups/forage_fk_abun.csv")
    forage_abun = forage_fk_abun %>%
      group_by(YEAR) %>%
      summarize(
        abundance_mean = mean(abundance),
        abundance_n = length(abundance),
        abundance_sd = sd(abundance),
        abundance_se = abundance_sd / sqrt(abundance_n),
        abundance_min = min(abundance),
        abundance_max = max(abundance))
    write_csv(forage_abun, "prep/data/rvc_trophic_groups/forage_abun.csv")
    
    forage_fk_den = read_csv("prep/data/rvc_trophic_groups/forage_fk_den.csv")
    forage_den = forage_fk_den %>%
      group_by(YEAR) %>%
      summarize(
        density_mean = mean(density),
        density_n = length(density),
        density_sd = sd(density),
        density_se = density_sd / sqrt(density_n),
        density_min = min(density),
        density_max = max(density))
    write_csv(forage_den, "prep/data/rvc_trophic_groups/forage_den.csv")
    
    forage_fk_bio = read_csv("prep/data/rvc_trophic_groups/forage_fk_bio_merged.csv")
    forage_bio = forage_fk_bio %>%
      group_by(YEAR) %>%
      summarize(
        biomass_mean = mean(biomass),
        biomass_n = length(biomass),
        biomass_sd = sd(biomass),
        biomass_se = biomass_sd / sqrt(biomass_n),
        biomass_min = min(biomass),
        biomass_max = max(biomass))
    write_csv(forage_bio, "prep/data/rvc_trophic_groups/forage_bio.csv")
    
    forage_fk_diveristy = read_csv("prep/data/rvc_trophic_groups/forage_fk_diversity.csv") 
    forage_div = forage_fk_diveristy %>%
      filter(protected_status != "0") %>%
      filter(protected_status != "1") %>%
      select(YEAR, richness, shannon, simpson)
    write_csv(forage_div, "prep/data/rvc_trophic_groups/forage_div.csv")
  }
}
```

- grouper
- snapper 
- hogfish
- seabass and hamlets 

### Grouper

| Scientific name               | Common name         | RVC code  |  
|:------------------------------|:--------------------|:----------|
|*Cephalopholis cruentata*      |	graysby	            | CEP CRUE  |
|*Cephalopholis fulva*          |	coney	              | CEP FULV  |
|*Dermatolepis inermis*         |	marbled grouper	    | DER INER  |
|*Epinephelus flavolimbatus*    |	yellowedge grouper	| EPI FLAV  |
|*Epinephelus itajara*          |	goliath grouper	    | EPI ITAJ  |
|*Epinephelus morio*            |	red grouper	        | EPI MORI  |
|*Epinephelus niveatus*	        | snowy grouper	      | EPI NIVE  |
|*Epinephelus striatus*         |	nassau grouper	    | EPI STRI  |
|*grouper-sea bass species*     |	grouper-sea bass sp |	SRR SPE.  |
|*Mycteroperca acutirostris*    |	western comb grouper| MYC ACUT  |
|*Mycteroperca bonaci*          |	black grouper	      | MYC BONA  |
|*Mycteroperca interstitialis*  |	yellowmouth grouper	| MYC INTE  |
|*Mycteroperca microlepis*      |	gag	                | MYC MICR  |
|*Mycteroperca phenax*          |	scamp	              | MYC PHEN  |
|*Mycteroperca tigris*          |	tiger grouper	      | MYC TIGR  |
|*Mycteroperca venenosa*        |	yellowfin grouper	  | MYC VENE  |

### Snappers

| Scientific name               | Common name         | RVC code  |  
|:------------------------------|:--------------------|:----------|
|*Lutjanus analis*              |	mutton snapper	    | LUT ANAL  |
|*Lutjanus apodus*	            | schoolmaster	      | LUT APOD  |
|*Lutjanus buccanella*	        | blackfin snapper	  | LUT BUCC  |
|*Lutjanus campechanus*	        | red snapper	        | Lut CAMP  |
|*Lutjanus cyanopterus*	        | cubera snapper	    | LUT CYAN  |
|*Lutjanus griseus*	            | gray snapper	      | LUT GRIS  |
|*Lutjanus jocu*	              | dog snapper	        | LUT JOCU  |
|*Lutjanus mahogoni*	          | mahogany snapper	  | LUT MAHO  |
|*Lutjanus synagris*	          | lane snapper	      | LUT SYNA  |
|*Ocyurus chrysurus*	          | yellowtail snapper	| OCY CHRY  |
|*Rhomboplites aurorubens*	    | vermilion snapper	  | RHO AURO  |
|*snapper species*	            | snapper species	    | LUT SPE.  |

### Hogfish

| Scientific name               | Common name         | RVC code  |  
|:------------------------------|:--------------------|:----------|
|*Bodianus pulchellus*	        | spotfin hogfish	    | BOD PULC  |
|*Bodianus rufus*	              | Spanish hogfish	    | BOD RUFU  |
|*Lachnolaimus maximus*	        | hogfish	            | LAC MAXI  |

### Sea bass and Hamlets

| Scientific name               | Common name           | RVC code  |  
|:------------------------------|:----------------------|:----------|
|*Alphestes afer*	              | mutton hamlet	        | ALP AFER  |
|*Centropristis ocyurus*	      | bank sea bass	        | CEN OCYU  |
|*Centropristis striata*        |	black sea bass	      | CEN STRI  |
|*Diplectrum formosum*          |	sand perch	          | DIP FORM  |
|*Epinephelus adscensionis*     |	rock hind	            | EPI ADSC  |
|*Epinephelus drummondhayi*     |	speckled hind	        | EPI DRUM  |
|*Epinephelus guttatus*         |	red hind	            | EPI GUTT  |
|*Hypoplectrus chlorurus*	      | yellowtail hamlet	    | HYP CHLO  |
|*Hypoplectrus gemma*	          | blue hamlet	          | HYP GEMM  |
|*Hypoplectrus guttavarius*	    | shy hamlet	          | HYP GUTT  |
|*Hypoplectrus hybrid*          |	hybrid hamlet	        | HYP HYBR  |
|*Hypoplectrus indigo*          |	indigo hamlet	        | HYP INDI  |
|*Hypoplectrus nigricans*       |	black hamlet	        | HYP NIGR  |
|*Hypoplectrus puella*          |	barred hamlet	        | HYP PUEL  |
|*Hypoplectrus sp.*             |	hamlet species	      | HYP SPE.  |
|*Hypoplectrus tann*            |	tan hamlet	          | HYP TANN  |
|*Hypoplectrus unicolor*        |	butter hamlet	        | HYP UNIC  |
|*Liopropoma eukrines*          |	wrasse basslet	      | LIO EUKR  |
|*Liopropoma mowbrayi*          |	cave basslet	        | LIO MOWB  |
|*Liopropoma rubre*	            | peppermint basslet	  | LIO RUBE  |
|*Paranthias furcifer*	        | atlantic creolefish	  | PAR FURC  |
|*Pristipomoides aquilonaris*	  | wenchman	            | PRI AQUI  |
|*Rypticus maculatus*	          | whitespotted soapfish	| RYP MACU  |
|*Rypticus saponaceus*	        | greater soapfish	    | RYP SAPO  |
|*Schultzea beta*	              | school bass	          | SCH BETA  |
|*Serranus annularis*	          | orangeback bass	      | SER ANNU  |
|*Serranus baldwini*	          | lantern bass	        | SER BALD  |
|*Serranus subligarius*	        | belted sandfish	      | SER SUBL  |
|*Serranus tabacarius*	        | tobaccofish	          | SER TABA  |
|*Serranus tigrinus*	          | harlequin bass	      | SER TIGR  |
|*Serranus tortugarum*	        | chalk bass	          | SER TORT  |

```{r grouper-snapper complex}

if(!all(file.exists("prep/data/rvc_trophic_groups/grouper_snapper_abun.csv",
                    "prep/data/rvc_trophic_groups/grouper_snapper_den.csv",
                    "prep/data/rvc_trophic_groups/grouper_snapper_bio.csv"))){
  
  grouper_fk_abun_csv       = "prep/data/rvc_trophic_groups/grouper_fk_abun.csv"
  grouper_fk_den_csv        = "prep/data/rvc_trophic_groups/grouper_fk_den.csv"
  grouper_fk_bio_csv        = "prep/data/rvc_trophic_groups/grouper_fk_bio_merged.csv"
  grouper_fk_diversity_csv  = "prep/data/rvc_trophic_groups/grouper_fk_diversity.csv"
  hogfish_fk_abun_csv       = "prep/data/rvc_trophic_groups/hogfish_fk_abun.csv"
  hogfish_fk_den_csv        = "prep/data/rvc_trophic_groups/hogfish_fk_den.csv"
  hogfish_fk_bio_csv        = "prep/data/rvc_trophic_groups/hogfish_fk_bio_merged.csv"
  seabass_fk_abun_csv       = "prep/data/rvc_trophic_groups/seabass_fk_abun.csv"
  seabass_fk_den_csv        = "prep/data/rvc_trophic_groups/seabass_fk_den.csv"
  seabass_fk_bio_csv        = "prep/data/rvc_trophic_groups/seabass_fk_bio_merged.csv"
  seabass_fk_diversity_csv  = "prep/data/rvc_trophic_groups/seabass_fk_diversity.csv"
  snapper_fk_abun_csv       = "prep/data/rvc_trophic_groups/snapper_fk_abun.csv"
  snapper_fk_den_csv        = "prep/data/rvc_trophic_groups/snapper_fk_den.csv"
  snapper_fk_bio_csv        = "prep/data/rvc_trophic_groups/snapper_fk_bio_merged.csv"
  snapper_fk_diversity_csv  = "prep/data/rvc_trophic_groups/snapper_fk_diversity.csv"
  
  if (!all(file.exists(grouper_fk_abun_csv,
                       grouper_fk_den_csv,grouper_fk_bio_csv,grouper_fk_diversity_csv,
                       hogfish_fk_abun_csv,hogfish_fk_den_csv,hogfish_fk_bio_csv,
                       seabass_fk_abun_csv,seabass_fk_den_csv,seabass_fk_bio_csv,seabass_fk_diversity_csv,
                       snapper_fk_abun_csv,snapper_fk_den_csv,snapper_fk_bio_csv,snapper_fk_diversity_csv))){
    
  grouper_spp_list= c("CEP CRUE", "MYC BONA", "EPI MORI", "EPI STRI", "EPI ITAJ", "CEP FULV", "MYC MICR", "MYC PHEN", "MYC VENE", "MYC INTE", "MYC TIGR", "EPI FLAV", "DER INER", "SRR SPE.", "Epi nive", "Myc acut")
  #abundance
  grouper_fk_abun = getDomainAbundance(RVCdata_FK, grouper_spp_list)
  write_csv(grouper_fk_abun, "prep/data/rvc_trophic_groups/grouper_fk_abun.csv")
  
  #density
  grouper_fk_den = getDomainDensity(RVCdata_FK, grouper_spp_list)
  write_csv(grouper_fk_den, "prep/data/rvc_trophic_groups/grouper_fk_den.csv")
  
  #biomass
  grouper_fk_bio = getDomainBiomass(RVCdata_FK, grouper_spp_list, RVCdata_FK$taxonomic_data)
  write_csv(grouper_fk_bio, "prep/data/rvc_trophic_groups/grouper_fk_bio_merged.csv")
  
  #diversity metrics 
  grouper_fk_abundance = grouper_fk_abun %>% 
    select(YEAR, protected_status, SPECIES_CD, abundance) %>%  
    group_by(YEAR, protected_status) %>% 
    nest(-YEAR) %>%  
    mutate(
      data_wide = map(data, ~ spread(data=.x, SPECIES_CD, abundance, fill =0)), 
      #species richness 
      richness = map(
        data_wide, 
        function(x) specnumber(x)),
      #simpson diversity as effective number of species 
      simpson = map( 
        data_wide,
        function(x) 1/(1 - diversity(x, index = 'simpson'))),
      #shannon diversity as effective number of species 
      shannon = map(
        data_wide,
        function(x) exp(diversity(x, index = 'shannon')))) %>%
    unnest(richness, simpson, shannon)
  
  grouper_fk_diversity = grouper_fk_abundance %>%
    select(YEAR, protected_status, richness, shannon, simpson) 
  write_csv(grouper_fk_diversity, "prep/data/rvc_trophic_groups/grouper_fk_diversity.csv")
  
  hogfish_spp_list = c("LAC MAXI", "BOD RUFU", "BOD PULC")
  
  #abundance
  hogfish_fk_abun = getDomainAbundance(RVCdata_FK, hogfish_spp_list, merge_protected = F)
  write_csv(hogfish_fk_abun, "prep/data/rvc_trophic_groups/hogfish_fk_abun.csv")
  
  #density
  hogfish_fk_den = getDomainDensity(RVCdata_FK, hogfish_spp_list, merge_protected = F)
  write_csv(hogfish_fk_den, "prep/data/rvc_trophic_groups/hogfish_fk_den.csv")
  
  #biomass
  hogfish_fk_bio = getDomainBiomass(RVCdata_FK, hogfish_spp_list, RVCdata_FK$taxonomic_data)
  write_csv(hogfish_fk_bio, "prep/data/rvc_trophic_groups/hogfish_fk_bio_merged.csv")
  
  #FK diversity
  hogfish_fk_abundance = hogfish_fk_abun %>% 
    select(YEAR, protected_status, SPECIES_CD, abundance) %>%  
    group_by(YEAR, protected_status) %>% #48 groups 
    nest(-YEAR) %>%  
    mutate(
      data_wide = map(data, ~ spread(data=.x, SPECIES_CD, abundance, fill = 0)), 
      #species richness 
      richness = map(
        data_wide, 
        function(x) specnumber(x)),
      #simpson diversity as effective number of species 
      simpson = map( 
        data_wide,
        function(x) 1/(1 - diversity(x, index = 'simpson'))),
      #shannon diversity as effective number of species 
      shannon = map(
        data_wide,
        function(x) exp(diversity(x, index = 'shannon')))) %>%
    unnest(richness, simpson, shannon)
  
  hogfish_fk_diversity = hogfish_fk_abundance %>%
    select(YEAR, protected_status, richness, simpson, shannon) 
  write_csv(hogfish_fk_diversity, "prep/data/rvc_trophic_groups/hogfish_fk_diversity.csv")
  
  sea_bass_hamlet_spp_list = c("SER TIGR","HYP GEMM","HYP UNIC","SER TABA","SER BALD","HYP PUEL","EPI ADSC","SER TORT","EPI GUTT","HYP NIGR","DIP FORM","HYP SPE.","SCH BETA","HYP TANN", "RYP SAPO","PAR FURC","HYP INDI","HYP GUTT","ALP AFER","LIO EUKR","SER ANNU","LIO RUBE","HYP HYBR","HYP CHLO","PRI AQUI","RYP MACU","SER SUBL","CEN OCYU","CEN STRI","Epi drum","Lio mowb")
  
  #abundance
  seabass_fk_abun = getDomainAbundance(RVCdata_FK, sea_bass_hamlet_spp_list, merge_protected = F)
  write_csv(seabass_fk_abun, "prep/data/rvc_trophic_groups/seabass_fk_abun.csv")
  
  #density
  seabass_fk_den = getDomainDensity(RVCdata_FK, sea_bass_hamlet_spp_list, merge_protected = F)
  write_csv(seabass_fk_den, "prep/data/rvc_trophic_groups/seabass_fk_den.csv")
  
  #biomass
  seabass_fk_bio = getDomainBiomass(RVCdata_FK, sea_bass_hamlet_spp_list,RVCdata_FK$taxonomic_data)
  write_csv(seabass_fk_bio, "prep/data/rvc_trophic_groups/seabass_fk_bio_merged.csv")
  
  #diversity
  seabass_fk_abundance = seabass_fk_abun %>% 
    select(YEAR, protected_status, SPECIES_CD, abundance) %>%  
    group_by(YEAR, protected_status) %>% #48 groups 
    nest(-YEAR) %>%  
    mutate(
      data_wide = map(data, ~ spread(data=.x, SPECIES_CD, abundance, fill = 0)), 
      #species richness 
      richness = map(
        data_wide, 
        function(x) specnumber(x)),
      #simpson diversity as effective number of species 
      simpson = map( 
        data_wide,
        function(x) 1/(1 - diversity(x, index = 'simpson'))),
      #shannon diversity as effective number of species 
      shannon = map(
        data_wide,
        function(x) exp(diversity(x, index = 'shannon')))) %>%
    unnest(richness, simpson, shannon)
  
  seabass_fk_diversity = seabass_fk_abundance %>%
    select(YEAR, protected_status, richness, simpson, shannon) 
  write_csv(seabass_fk_diversity, "prep/data/rvc_trophic_groups/seabass_fk_diversity.csv")
  
  snapper_spp_list = c("OCY CHRY", "LUT GRIS", "LUT ANAL", "LUT APOD", "LUT SYNA", "LUT MAHO", "LUT JOCU", "LUT BUCC", 'LUT SPE.', "LUT CYAN", "RHO AURO", "Lut camp")
  
  #abundance
  snapper_fk_abun = getDomainAbundance(RVCdata_FK, snapper_spp_list, merge_protected = F)
  write_csv(snapper_fk_abun, "prep/data/rvc_trophic_groups/snapper_fk_abun.csv")
  
  #density
  snapper_fk_den = getDomainDensity(RVCdata_FK, snapper_spp_list, merge_protected = F)
  write_csv(snapper_fk_den, "prep/data/rvc_trophic_groups/snapper_fk_den.csv")
  
  #biomass
  snapper_fk_bio = getDomainBiomass(RVCdata_FK, snapper_spp_list, RVCdata_FK$taxonomic_data)
  write_csv(snapper_fk_bio, "prep/data/rvc_trophic_groups/snapper_fk_bio_merged.csv")
  
  #diversity
  snapper_fk_abundance = snapper_fk_abun %>% 
    select(YEAR, protected_status, SPECIES_CD, abundance) %>%  
    group_by(YEAR, protected_status) %>% #48 groups 
    nest(-YEAR) %>%  
    mutate(
      data_wide = map(data, ~ spread(data=.x, SPECIES_CD, abundance, fill = 0)), 
      #species richness 
      richness = map(
        data_wide, 
        function(x) specnumber(x)),
      #simpson diversity as effective number of species 
      simpson = map( 
        data_wide,
        function(x) 1/(1 - diversity(x, index = 'simpson'))),
      #shannon diversity as effective number of species 
      shannon = map(
        data_wide,
        function(x) exp(diversity(x, index = 'shannon')))) %>%
    unnest(richness, simpson, shannon)
  
  snapper_fk_diversity = snapper_fk_abundance %>%
    select(YEAR, protected_status, richness, simpson, shannon) 
  write_csv(snapper_fk_diversity, "prep/data/rvc_trophic_groups/snapper_fk_diversity.csv")
  
  }
  
  #abundance data 
  snapper_abun = read_csv("prep/data/rvc_trophic_groups/snapper_fk_abun.csv")
  seabass_abun = read_csv("prep/data/rvc_trophic_groups/seabass_fk_abun.csv")
  hogfish_abun = read_csv("prep/data/rvc_trophic_groups/hogfish_fk_abun.csv")
  grouper_abun = read_csv("prep/data/rvc_trophic_groups/grouper_fk_abun.csv")
  
  snapper_grouper_abun = rbind(snapper_abun,seabass_abun,hogfish_abun,grouper_abun) #1999-2016 
  
  snapper_grouper_abun = snapper_grouper_abun %>%
  group_by(YEAR) %>%
      summarize(
        abundance_mean = mean(abundance),
        abundance_n = length(abundance),
        abundance_sd = sd(abundance),
        abundance_se = abundance_sd / sqrt(abundance_n),
        abundance_min = min(abundance),
        abundance_max = max(abundance))
    write_csv(snapper_grouper_abun, "prep/data/rvc_trophic_groups/snapper_grouper_abun.csv")
    
  #biomass data 
  snapper_bio = read_csv("prep/data/rvc_trophic_groups/snapper_fk_bio_merged.csv")
  seabass_bio = read_csv("prep/data/rvc_trophic_groups/seabass_fk_bio_merged.csv")
  hogfish_bio = read_csv("prep/data/rvc_trophic_groups/hogfish_fk_bio_merged.csv")
  grouper_bio = read_csv("prep/data/rvc_trophic_groups/grouper_fk_bio_merged.csv")
  
  snapper_grouper_bio = rbind(snapper_bio,seabass_bio,hogfish_bio,grouper_bio) #1999-2016 
  
  snapper_grouper_bio = snapper_grouper_bio %>%
      group_by(YEAR) %>%
      summarize(
        biomass_mean = mean(biomass),
        biomass_n = length(biomass),
        biomass_sd = sd(biomass),
        biomass_se = biomass_sd / sqrt(biomass_n),
        biomass_min = min(biomass),
        biomass_max = max(biomass))
    write_csv(snapper_grouper_bio, "prep/data/rvc_trophic_groups/snapper_grouper_bio.csv")

  #density data 
  snapper_den = read_csv("prep/data/rvc_trophic_groups/snapper_fk_den.csv")
  seabass_den = read_csv("prep/data/rvc_trophic_groups/seabass_fk_den.csv")
  hogfish_den = read_csv("prep/data/rvc_trophic_groups/hogfish_fk_den.csv")
  grouper_den = read_csv("prep/data/rvc_trophic_groups/grouper_fk_den.csv")
  
  snapper_grouper_den = rbind(snapper_den,seabass_den,hogfish_den,grouper_den) #1999-2016 
  
  snapper_grouper_den = snapper_grouper_den %>%
      group_by(YEAR) %>%
      summarize(
        density_mean = mean(density),
        density_n = length(density),
        density_sd = sd(density),
        density_se = density_sd / sqrt(density_n),
        density_min = min(density),
        density_max = max(density))
    write_csv(snapper_grouper_den, "prep/data/rvc_trophic_groups/snapper_grouper_den.csv")
    
  #diversity data 
  snapper_diversity = read_csv("prep/data/rvc_trophic_groups/snapper_fk_diversity.csv")
  seabass_diversity = read_csv("prep/data/rvc_trophic_groups/seabass_fk_diversity.csv")
  hogfish_diversity = read_csv("prep/data/rvc_trophic_groups/hogfish_fk_diversity.csv")
  grouper_diversity = read_csv("prep/data/rvc_trophic_groups/grouper_fk_diversity.csv")
  
  snapper_grouper_diversity = rbind(snapper_diversity,seabass_diversity,hogfish_diversity,grouper_diversity) #1999-2016 
  
  snapper_grouper_diversity = snapper_grouper_diversity %>%
    filter(protected_status != "0") %>%
    filter(protected_status != "1") %>%
    group_by(YEAR) %>%
    summarize(
      richness_mean = mean(richness),
      richness_n = length(richness),
      richness_sd = sd(richness),
      richness_se = richness_sd / sqrt(richness_n),
      richness_min = min(richness),
      richness_max = max(richness),
      simpson_mean = mean(simpson),
      simpson_n = length(simpson),
      simpson_sd = sd(simpson),
      simpson_se = simpson_sd / sqrt(simpson_n),
      simpson_min = min(simpson),
      simpson_max = max(simpson),
      shannon_mean = mean(shannon),
      shannon_n = length(shannon),
      shannon_sd = sd(shannon),
      shannon_se = shannon_sd / sqrt(shannon_n),
      shannon_min = min(shannon),
      shannon_max = max(shannon))
  
    write_csv(snapper_grouper_diversity, "prep/data/rvc_trophic_groups/snapper_grouper_diversity.csv")  
}
  
```

### Parrotfish

| Scientific name               | Common name           | RVC code  |  
|:------------------------------|:----------------------|:----------|
|*Cryptotomus roseus*	          | bluelip parrotfish	  | CRY ROSE  |
|*Nicholsina usta*	            | emerald parrotfish	  | NIC USTA  |
|*Scarus coelestinus*	          | midnight parrotfish	  | SCA COEL  |
|*Scarus coeruleus*	            | blue parrotfish	      | SCA COER  |
|*Scarus guacamaia*	            | rainbow parrotfish	  | SCA GUAC  |
|*Scarus iseri*	                | striped parrotfish	  | SCA ISER  |
|*Scarus sp.*	                  | parrotfish species	  | SCA SPE.  |
|*Scarus taeniopterus*          |	princess parrotfish	  | SCA TAEN  |
|*Scarus vetula*	              | queen parrotfish	    | SCA VETU  |
|*Sparisoma atomarium*	        | greenblotch parrotfish|	SPA ATOM  |
|*Sparisoma aurofrenatum*	      | redband parrotfish	  | SPA AURO  |
|*Sparisoma chrysopterum*	      | redtail parrotfish	  | SPA CHRY  |
|*Sparisoma radians*	          | bucktooth parrotfish	| SPA RADI  |
|*Sparisoma rubripinne*	        | yellowtail parrotfish	| SPA RUBR  |
|*Sparisoma sp.*	              | parrotfish species	  | SPA SPE.  |
|*Sparisoma viride*	            | stoplight parrotfish	| SPA VIRI  |

```{r parrotfish}

if(!all(file.exists("prep/data/rvc_trophic_groups/parrotfish_abun.csv",
                    "prep/data/rvc_trophic_groups/parrotfish_den.csv",
                    "prep/data/rvc_trophic_groups/parrotfish_bio.csv",
                    "prep/data/rvc_trophic_groups/parrotfish_div.csv"))){
  
  parrotfish_fk_abun_csv     = "prep/data/parrotfish_fk_abun.csv"        #abundance
  parrotfish_fk_den_csv      = "prep/data/parrotfish_fk_den.csv"           #density
  parrotfish_fk_bio_csv      = "prep/data/parrotfish_fk_bio_merged.csv"    #biomass 
  parrotfish_fk_diversity_csv= "prep/data/parrotfish_fk_diversity.csv"   #diversity 
  
  }else{
    if (!all(file.exists(
    parrotfish_fk_abun_csv, 
    parrotfish_fk_den_csv, 
    parrotfish_fk_bio_csv,
    parrotfish_fk_diversity_csv))){
    
    parrotfish_spp_list = c('SCA ISER', "SCA COEL", "SPA AURO", "SPA VIRI", "SPA ATOM", "SCA TAEN", 'SPA RUBR', "SPA CHRY", "SCA VETU", "SCA COER", "SCA GUAC", "CRY ROSE", "SPA RADI", "SCA SPE.", "NIC USTA", "SPA SPE.")
  
    #abundance
    parrotfish_fk_abun = getDomainAbundance(RVCdata_FK, parrotfish_spp_list, merge_protected = F)
    write_csv(parrotfish_fk_abun, "prep/data/rvc_trophic_groups/parrotfish_fk_abun.csv")
  
    #density
    parrotfish_fk_den = getDomainDensity(RVCdata_FK, parrotfish_spp_list, merge_protected = F)
    write_csv(parrotfish_fk_den, "prep/data/rvc_trophic_groups/parrotfish_fk_den.csv")
  
    #biomass
    parrotfish_fk_bio = getDomainBiomass(RVCdata_FK, parrotfish_spp_list, RVCdata_FK$taxonomic_data)
    write_csv(parrotfish_fk_bio, "prep/data/rvc_trophic_groups/parrotfish_fk_bio_merged.csv")
  
    #diversity
    parrotfish_fk_abundance = parrotfish_fk_abun %>% 
      select(YEAR, protected_status, SPECIES_CD, abundance) %>%  
      group_by(YEAR, protected_status) %>% #48 groups 
      nest(-YEAR) %>%  
      mutate(
        data_wide = map(data, ~ spread(data=.x, SPECIES_CD, abundance, fill = 0)), 
        #species richness 
        richness = map(
          data_wide, 
          function(x) specnumber(x)),
        #simpson diversity as effective number of species 
        simpson = map( 
          data_wide,
          function(x) 1/(1 - diversity(x, index = 'simpson'))),
        #shannon diversity as effective number of species 
        shannon = map(
          data_wide,
          function(x) exp(diversity(x, index = 'shannon')))) %>%
      unnest(richness, simpson, shannon)
  
    parrotfish_fk_diversity = parrotfish_fk_abundance %>%
      select(YEAR, protected_status, richness, simpson, shannon) 
    write_csv(parrotfish_fk_diversity, "prep/data/rvc_trophic_groups/parrotfish_fk_diversity.csv")
  
    }else{
    
    parrotfish_abun = read_csv("prep/data/rvc_trophic_groups/parrotfish_fk_abun.csv")
    parrotfish_abun = parrotfish_abun %>%
      group_by(YEAR) %>%
      summarize(
        abundance_mean = mean(abundance),
        abundance_n = length(abundance),
        abundance_sd = sd(abundance),
        abundance_se = abundance_sd / sqrt(abundance_n),
        abundance_min = min(abundance),
        abundance_max = max(abundance))
    write_csv(parrotfish_abun, "prep/data/rvc_trophic_groups/parrotfish_abun.csv")
    
    parrotfish_den = read_csv("prep/data/rvc_trophic_groups/parrotfish_fk_den.csv")
    parrotfish_den = parrotfish_den %>%
      group_by(YEAR) %>%
      summarize(
        density_mean = mean(density),
        density_n = length(density),
        density_sd = sd(density),
        density_se = density_sd / sqrt(density_n),
        density_min = min(density),
        density_max = max(density))
    write_csv(parrotfish_den, "prep/data/rvc_trophic_groups/parrotfish_den.csv")
    
    parrotfish_bio = read_csv("prep/data/rvc_trophic_groups/parrotfish_fk_bio_merged.csv")
    parrotfish_bio = parrotfish_bio %>%
      group_by(YEAR) %>%
      summarize(
        biomass_mean = mean(biomass),
        biomass_n = length(biomass),
        biomass_sd = sd(biomass),
        biomass_se = biomass_sd / sqrt(biomass_n),
        biomass_min = min(biomass),
        biomass_max = max(biomass))
    write_csv(parrotfish_bio, "prep/data/rvc_trophic_groups/parrotfish_bio.csv")
    
    parrotfish_diveristy = read_csv("prep/data/rvc_trophic_groups/parrotfish_fk_diversity.csv") 
    parrotfish_div = parrotfish_diveristy %>%
      filter(protected_status != "0") %>%
      filter(protected_status != "1") %>%
      select(YEAR, richness, shannon, simpson)
    write_csv(parrotfish_div, "prep/data/rvc_trophic_groups/parrotfish_div.csv")
}

```

### Jacks/Opportunists 

| Scientific name               | Common name         | RVC code  |  
|:------------------------------|:--------------------|:----------|
|*Alectis ciliaris*	            | african pompano	    | ALE CILI  |
|*Caranx crysos*	              | blue runner	        | CAR CRYS  |
|*Caranx bartholomaei*          |	yellow jack	        | CAR BART  |
|*Caranx latus*                 |	horse-eye jack	    | CAR LATU  |
|*Caranx lugubris*              |	black jack	        | CAR LUGU  |
|*Caranx ruber*	                | bar jack	          | CAR RUBE  |
|*Caranx sp.*	                  | jack species	      | CAR SPE.  |
|*Decapterus punctatus*	        | round scad	        | DEC PUNC  |
|*Decapterus macarellus*	      | mackerel scad	      | DEC MACA  |
|*Elagatis bipinnulata*	        | rainbow runner	    | ELA BIPI  |
|*Euthynnus alletteratus*	      | little tunny	      | EUT ALLE  |
|*Oligoplites saurus*           |	leatherjack	        | OLI SAUR  |
|*Sphyraena barracuda*	        | great barracuda	    | SPH BARR  |
|*Rachycentron canadum*	        | cobia	              | RAC CANA  |
|*Scomberomorus regalis*	      | cero	              | SCO REGA  |
|*Sphyraena picudilla*	        | southern sennet	    | SPH PICU  |
|*Selene vomer*	                | lookdown	          | SEL VOME  |
|*Seriola rivoliana*	          | almaco jack       	| SER RIVO  |
|*Scomberomorus maculatus*	    | spanish mackerel  	| SCO MACU  |
|*Seriola dumerili*	            | greater amberjack 	| SER DUME  |
|*Scomberomorus cavalla*	      | king mackerel	      | SCO CAVA  |
|*Sphyraena guachancho*	        | guaguanche	        | SPH GUAC  |
|*Seriola sp.*	                | jack species        |	SER SPE.  |
|*Seriola zonata*	              | banded rudderfish	  | SER ZONA  |
|*Trachinotus falcatus*	        | permit	            | TRA FALC  |

```{r jacks}

if(!all(file.exists("prep/data/rvc_trophic_groups/jacks_abun.csv",
                    "prep/data/rvc_trophic_groups/jacks_den.csv",
                    "prep/data/rvc_trophic_groups/jacks_bio.csv",
                    "prep/data/rvc_trophic_groups/jacks_div.csv"))){
  
  opportunists_fk_abun_csv      = "prep/data/rvc_trophic_groups/opportunists_fk_abun.csv"
  opportunists_fk_den_csv       = "prep/data/rvc_trophic_groups/opportunists_fk_den.csv"
  opportunists_fk_bio_csv       = "prep/data/rvc_trophic_groups/opportunists_fk_den_merged.csv"
  opportunists_fk_diversity_csv = "prep/data/rvc_trophic_groups/opportunists_fk_diversity.csv"
  
  if(!all(file.exists(opportunists_fk_abun_csv,opportunists_fk_den_csv,opportunists_fk_bio_csv,opportunists_fk_diversity_csv))){
    
    opportunists_spp_list = c("CAR RUBE", "SPH BARR", "ALE CILI","DEC PUNC", "DEC MACA", "CAR CRYS", "CAR BART", "SCO REGA","SPH PICU", "CAR LATU", "SEL VOME", "ELA BIPI", "CAR SPE.", "TRA FALC", "EUT ALLE", "SER RIVO", "SCO MACU", "SER DUME", "SCO CAVA", "SPH GUAC", "CAR LUGU", "OLI SAUR", "SER SPE.", "Rac cana", "Ser zona")
    
    #abundance
    opportunists_fk_abun = getDomainAbundance(RVCdata_FK, opportunists_spp_list, merge_protected = F)
    write_csv(opportunists_fk_abun, "prep/data/rvc_trophic_groups/opportunists_fk_abun.csv")
    
    #density
    opportunists_fk_den = getDomainDensity(RVCdata_FK, opportunists_spp_list, merge_protected = F)
    write_csv(opportunists_fk_den, "prep/data/rvc_trophic_groups/opportunists_fk_den.csv")
    
    #biomass
    opportunists_fk_bio = getDomainBiomass(RVCdata_FK, opportunists_spp_list, RVCdata_FK$taxonomic_data)
    write_csv(opportunists_fk_bio, "prep/data/rvc_trophic_groups/opportunists_fk_bio_merged.csv")
    
    #diversity
    opportunists_fk_abundance = opportunists_fk_abun %>% 
      select(YEAR, protected_status, SPECIES_CD, abundance) %>%  
      group_by(YEAR, protected_status) %>% #48 groups 
      nest(-YEAR) %>%  
      mutate(
        data_wide = map(data, ~ spread(data=.x, SPECIES_CD, abundance, fill = 0)), 
        #species richness 
        richness = map(
          data_wide, 
          function(x) specnumber(x)),
        #simpson diversity as effective number of species 
        simpson = map( 
          data_wide,
          function(x) 1/(1 - diversity(x, index = 'simpson'))),
        #shannon diversity as effective number of species 
        shannon = map(
          data_wide,
          function(x) exp(diversity(x, index = 'shannon')))) %>%
      unnest(richness, simpson, shannon)
    
    opportunists_fk_diversity = opportunists_fk_abundance %>%
      select(YEAR, protected_status, richness, simpson, shannon) 
    write_csv(opportunists_fk_diversity, "prep/data/rvc_trophic_groups/opportunists_fk_diversity.csv")
    
  }else{
    
    #abundance
    opportunists_abun = read_csv("prep/data/rvc_trophic_groups/hogfish_fk_abun.csv")
    
    opportunists_abun = opportunists_fk_abun %>%
      group_by(YEAR) %>%
      summarize(
        abundance_mean = mean(abundance),
        abundance_n = length(abundance),
        abundance_sd = sd(abundance),
        abundance_se = abundance_sd / sqrt(abundance_n),
        abundance_min = min(abundance),
        abundance_max = max(abundance))
    write_csv(opportunists_abun, "prep/data/rvc_trophic_groups/opportunists_abun.csv")
    
    #biomass
    opportunists_bio = read_csv("prep/data/rvc_trophic_groups/opportunists_fk_bio_merged.csv")
    
    opportunists_biomass = opportunists_bio %>%
      group_by(YEAR) %>%
      summarize(
        biomass_mean = mean(biomass),
        biomass_n = length(biomass),
        biomass_sd = sd(biomass),
        biomass_se = biomass_sd / sqrt(biomass_n),
        biomass_min = min(biomass),
        biomass_max = max(biomass))
    write_csv(opportunists_biomass, "prep/data/rvc_trophic_groups/opportunists_bio.csv")
    
    
    #density
    opportunists_den = read_csv("prep/data/rvc_trophic_groups/opportunists_fk_den.csv")
    
    opportunists_density = opportunists_den %>%
      group_by(YEAR) %>%
      summarize(
        density_mean = mean(density),
        density_n = length(density),
        density_sd = sd(density),
        density_se = density_sd / sqrt(density_n),
        density_min = min(density),
        density_max = max(density))
    write_csv(opportunists_density, "prep/data/rvc_trophic_groups/opportunists_den.csv")
    
    #diversity
    opportunists_diversity = read_csv("prep/data/rvc_trophic_groups/opportunists_fk_diversity.csv")
    
    opportunists_div = opportunists_diversity %>%
      filter(protected_status != "0") %>%
      filter(protected_status != "1") %>%
      select(YEAR, richness, shannon, simpson)
    write_csv(opportunists_div, "prep/data/rvc_trophic_groups/opportunists_div.csv")
  }
}

```

- grunt
- higher level reef fish
- small reef reef fish 

###Grunts 

| Scientific name               | Common name         | RVC code  |  
|:------------------------------|:--------------------|:----------|
|*Anisotremus surinamensis*	    | black margate	      | ANI SURI  |
|*Anisotremus virginicus*	      | porkfish	          | ANI VIRG  |
|*Haemulon album*	              | margate	            | HAE ALBU  |
|*Haemulon aurolineatum*	      | tomtate	            | HAE AURO  | 
|*Haemulon carbonarium*	        | caesar grunt	      | HAE CARB  |
|*Haemulon chrysargyreum*       |	smallmouth grunt	  | HAE CHRY  |
|*Haemulon flavolineatum*       |	french grunt	      | HAE FLAV  |
|*Haemulon macrostomum*	        | spanish grunt	      | HAE MACR  |
|*Haemulon melanurum*	          | cottonwick	        | HAE MELA  |
|*Haemulon parra*               |	sailors choice	    | HAE PARR  |
|*Haemulon plumierii*           |	white grunt	        | HAE PLUM  | 
|*Haemulon sciurus*             |	bluestriped grunt	  | HAE SCIU  |
|*Haemulon sp.*                 |	grunt species	      | HAE SPE.  |
|*Haemulon striatum*            |	striped grunt	      | HAE STRI  | 

### Small Reef Fish

| Scientific name               | Common name         | RVC code  |  
|:------------------------------|:--------------------|:----------|

### Higher Level Reef Fish

| Scientific name               | Common name         | RVC code  |  
|:------------------------------|:--------------------|:----------|

```{r grunt etc.}

if(!all(file.exists("prep/data/rvc_trophic_groups/grunt_etc_abun.csv",
                    "prep/data/rvc_trophic_groups/grunt_etc_den.csv",
                    "prep/data/rvc_trophic_groups/grunt_etc_bio.csv",
                    "prep/data/rvc_trophic_groups/grunt_etc_div.csv"))){
  
  grunt_fk_abun_csv             = "prep/data/rvc_trophic_groups/grunt_fk_abun.csv"
  grunt_fk_den_csv              = "prep/data/rvc_trophic_groups/grunt_fk_den.csv"
  grunt_fk_bio_csv              = "prep/data/rvc_trophic_groups/grunt_fk_bio_merged.csv"
  grunt_fk_diversity_csv        = "prep/data/rvc_trophic_groups/grunt_fk_diversity.csv"
  higher_reef_fk_abun_csv       = "prep/data/rvc_trophic_groups/higher_reef_fk_abun.csv"
  higher_reef_fk_den_csv        = "prep/data/rvc_trophic_groups/higher_reef_fk_den.csv"
  higher_reef_fk_bio_csv        = "prep/data/rvc_trophic_groups/higher_reef_fk_bio_merged.csv"
  higher_reef_fk_diversity_csv  = "prep/data/rvc_trophic_groups/higher_reef_fk_diversity.csv"
  small_reef_fk_abun_csv        = "prep/data/rvc_trophic_groups/small_reef_fk_abun.csv"
  small_reef_fk_den_csv         = "prep/data/rvc_trophic_groups/small_reef_fk_den.csv"
  small_reef_fk_bio_csv         = "prep/data/rvc_trophic_groups/small_reef_fk_bio_merged.csv"
  small_reef_fk_diversity_csv   = "prep/data/rvc_trophic_groups/small_reef_fk_diversity.csv"
  
  
  if(!all(file.exists(grunt_fk_abun_csv,grunt_fk_den_csv,grunt_fk_bio_csv,grunt_fk_diversity_csv,
                      higher_reef_fk_abun_csv,higher_reef_fk_den_csv,higher_reef_fk_bio_csv,higher_reef_fk_diversity_csv,
                      small_reef_fk_abun_csv,small_reef_fk_den_csv,small_reef_fk_bio_csv,small_reef_fk_diversity_csv))){
    
    grunt_spp_list = c("HAE SCIU", "HAE AURO", "HAE PLUM", "HAE SPE.", "HAE FLAV", "HAE CHRY", "ANI VIRG","HAE MELA", "HAE CARB", "HAE PARR", "HAE STRI", "HAE MACR", "ANI SURI", "HAE ALBU")
    
    #abundance
    grunt_fk_abun = getDomainAbundance(RVCdata_FK, grunt_spp_list)
    write_csv(grunt_fk_abun, "prep/data/rvc_trophic_groups/grunt_fk_abun.csv")
    
    #density
    grunt_fk_den = getDomainDensity(RVCdata_FK, grunt_spp_list)
    write_csv(grunt_fk_den, "prep/data/rvc_trophic_groups/grunt_fk_den.csv")
    
    #biomass
    grunt_fk_bio = getDomainBiomass(RVCdata_FK, grunt_spp_list, RVCdata_FK$taxonomic_data)
    write_csv(grunt_fk_bio, "prep/data/rvc_trophic_groups/grunt_fk_bio_merged.csv")
    
    #diversity metrics 
    grunt_fk_abundance = grunt_fk_abun %>% 
      select(YEAR, protected_status, SPECIES_CD, abundance) %>%  
      group_by(YEAR, protected_status) %>% #48 groups 
      nest(-YEAR) %>%  
      mutate(
        data_wide = map(data, ~ spread(data=.x, SPECIES_CD, abundance, fill = 0)), 
        #species richness 
        richness = map(
          data_wide, 
          function(x) specnumber(x)),
        #simpson diversity as effective number of species 
        simpson = map( 
          data_wide,
          function(x) 1/(1 - diversity(x, index = 'simpson'))),
        #shannon diversity as effective number of species 
        shannon = map(
          data_wide,
          function(x) exp(diversity(x, index = 'shannon')))) %>%
      unnest(richness, simpson, shannon)
    
    grunt_fk_diversity = grunt_fk_abundance %>%
      select(YEAR, protected_status, richness, simpson, shannon) 
    write_csv(grunt_fk_diversity, "prep/data/rvc_trophic_groups/grunt_fk_diversity.csv")
    
    higher_reef_spp_list = c("THA BIFA","STE PART","GYM MORI","PRI AREN","ABU SAXA","HAL BIVI","HAL GARN","HAL MACU","CHR MULT","CLE PARR","STE PLAN","CHR CYAN","MUL MART","MIC CHRY","STE ADUS","CHA CAPI","STE VARI","PEM SCHO","POM ARCU","CAN ROST","CHA OCEL","PSE MACU","HOL TRIC","CHR SCOT","HAL RADI","CAL CALA","CHA SEDE","XYR SPLE","STE LEUC","HOL CILI","POM PARU","AUL MACU","CHA STRI","OPI AURI","HAL POEY","CHR INSO","HOL ADSC","ODO DENT","STE DIEN","HOL BERM","CHA FABE","XYR MART","GER CINE","HOL RUFU","LAC TRIQ","PAR ACUM","ALU SCRI","BAL CAPR","MAL PLUM","CAN SUFF","CAN PULL","HET CRUE","OGC SPE.","XYR SPE.","ACA QUAD","CAL BAJO","MON TUCK","SPH SPEN","ARC RHOM","DIO HOLO","LAC BICA","XYR NOVA","BAL VETU","DIO HYST","HAL CYAN","GYM FUNE","DIP HOLB","APO PSEU","HOL TOWN","EQU PUNC","LAC TRIG","HAL PICT","MYR JACO","SAR VEXI","SYN INTE","HAL CAUD","AMB PINO","GYM MILI","ALU SCHO","ACA POLY","CAL PROR","CHI SCHO","HET LONG","SCO PLUM","CHR ENCH","CAL SPE.","APO MACU","CAN MACR","PAR UMBR","LAG RHOM","EQU LANC","CAL PENN","STE HISP","APO BINO","NEO MARI","ARC PROB","CHI ANTE","APO TOWN","DOR MEGA","REM REMO","OPI WHIT","CEN ARGI","SYN FOET","AST SPE.","CAR HIPP","SAR CORU","OPI MACR","GYM VICI","DIP ARGE","PAG PAGR","PAR BAIR","BOT LUNA","EUC ARGE","PRO ACUL","LAB GOBI","FIS TABA","ALU SPE.","GRA LORE","MON CILI","ECH NEUC","ALU MONO","BOT OCEL","BAL SPE.","PAR ALBI","STE SPE.","APO QUAD","ENC NIGR","GYM SAXI","OPI SPE.","ELO SAUR","MAN BIRO","APO AURO","SPH TEST","CAL NODO","DIO SPE.","SCO CARI","LAB FILA","AHL EGMO","ANT OCEL","AST GUTT","LAB SPE.","LAB SPE.","MUR RETI","AST STEL","EUC GULA","HOL SPE.","ORT CHRY","PRI OPHR","LAB BUCC","ACA SOLA","CAL LEUC","ENC CARY","GYM NIGR","MYR BREV","SAR BULL","TRA GOOD","TRA LATH","UMB CORO","EUC JONE","EUCLEFR","STY LATE","SYA MICR","SYN SCOV","APO PHEN","HIP REID","PHA XENU","AST PUNC","CHI ATIN","CHL CHRY","HIP EREC","OGC NASU","SER PHOE","SPH NEPH","MYR OCEL","PRI RUBI","RHI LENT","ALB VULP","DAC VOLI","RYP BIST","BRO BARB","HYP GUMM","OPS TAU_","EMM ATLA","UPE PARV","SYN SYNO")
    
    #abundance
    higher_reef_fk_abun = getDomainAbundance(RVCdata_FK, higher_reef_spp_list)
    write_csv(higher_reef_fk_abun, "prep/data/rvc_trophic_groups/higher_reef_fk_abun_merged.csv")
    
    #density
    higher_reef_fk_den = getDomainDensity(RVCdata_FK, higher_reef_spp_list)
    write_csv(higher_reef_fk_den, "prep/data/rvc_trophic_groups/higher_reef_fk_den_merged.csv")
    
    #biomass
    higher_reef_fk_bio = getDomainBiomass(RVCdata_FK, higher_reef_spp_list, RVCdata_FK$taxonomic_data)
    write_csv(higher_reef_fk_bio, "prep/data/rvc_trophic_groups/higher_reef_fk_bio_merged.csv")
    
    #fk diversity
    higher_reef_fk_abundance = higher_reef_fk_abun %>% 
      select(YEAR, protected_status, SPECIES_CD, abundance) %>%  
      group_by(YEAR, protected_status) %>% #48 groups 
      nest(-YEAR) %>%  
      mutate(
        data_wide = map(data, ~ spread(data=.x, SPECIES_CD, abundance, fill = 0)), 
        #species richness 
        richness = map(
          data_wide, 
          function(x) specnumber(x)),
        #simpson diversity as effective number of species 
        simpson = map( 
          data_wide,
          function(x) 1/(1 - diversity(x, index = 'simpson'))),
        #shannon diversity as effective number of species 
        shannon = map(
          data_wide,
          function(x) exp(diversity(x, index = 'shannon')))) %>%
      unnest(richness, simpson, shannon)
    
    higher_reef_diversity = higher_reef_fk_abundance %>%
      select(YEAR, protected_status, richness, simpson, shannon) 
    write_csv(higher_reef_diversity, "prep/data/rvc_trophic_groups/higher_reef_fk_diversity.csv")
    
    small_reef_fish_spp_list = c("COR PERS","ACA COER","KYP SECT","ACA BAHI","ACA CHIR","COR GLAU","ELA OCEA","GNA THOM","PTE CALL","ACA SPE.","MAL TRIA","COR DICR","PAR MARM","MAL MACR","OPH MACC","CTE SAEP","MIC CARR","PTE HELE","HEM SIMU","SCA CRIS","COR EIDO","GOB SPE.","TYL CROC","MIC MICR","ACA ASPE","COR SPE.","BLE SPE.","STR NOTA","MEL NIGE","HYP BERM","EMB PAND","ELA EVEL","ACA CHAP","EMB BAHA","COR LIPE","NES LONG","ENN BOEH","ACA MARI","LAB NUCH","MAL VERS","MAL GILL","ELA MACR","PRI HIPO","GOB DILE","ELA SAUC","ELA XANT","STR TIMU","MAL AURO","PAR MARN","CHA LIMB","OXY STIG","ELA HORS","PAR NIGR","COR PUNC","ACA SPIN","BOL BOQU","ElA RAND","LAB KALI","LAB NIGR","ENN ALTI")
    
    #abundance
    small_reef_fk_abun = getDomainAbundance(RVCdata_FK, small_reef_fish_spp_list, merge_protected = F)
    write_csv(small_reef_fk_abun, "prep/data/rvc_trophic_groups/small_reef_fk_abun.csv")
    
    #density
    small_reef_fk_den = getDomainDensity(RVCdata_FK, small_reef_fish_spp_list, merge_protected = F)
    write_csv(small_reef_fk_den, "prep/data/rvc_trophic_groups/small_reef_fk_den.csv")
    
    #biomass
    small_reef_fk_bio = getDomainBiomass(RVCdata_FK, small_reef_fish_spp_list, RVCdata_FK$taxonomic_data)
    write_csv(small_reef_fk_bio, "prep/data/rvc_trophic_groups/small_reef_fk_bio_merged.csv")
    
    #fk diversity
    small_reef_fk_abundance = small_reef_fk_abun %>% 
      select(YEAR, protected_status, SPECIES_CD, abundance) %>%  
      group_by(YEAR, protected_status) %>% #48 groups 
      nest(-YEAR) %>%  
      mutate(
        data_wide = map(data, ~ spread(data=.x, SPECIES_CD, abundance, fill = 0)), 
        #species richness 
        richness = map(
          data_wide, 
          function(x) specnumber(x)),
        #simpson diversity as effective number of species 
        simpson = map( 
          data_wide,
          function(x) 1/(1 - diversity(x, index = 'simpson'))),
        #shannon diversity as effective number of species 
        shannon = map(
          data_wide,
          function(x) exp(diversity(x, index = 'shannon')))) %>%
      unnest(richness, simpson, shannon)
    
    small_reef_fk_diversity = small_reef_fk_abundance %>%
      select(YEAR, protected_status, richness, simpson, shannon) 
    write_csv(small_reef_fk_diversity, "prep/data/rvc_trophic_groups/small_reef_fk_diversity.csv")
    
  }
  
  #abundance data 
  grunt_abun        = read_csv("prep/data/rvc_trophic_groups/snapper_fk_abun.csv")
  higher_reef_abun  = read_csv("prep/data/rvc_trophic_groups/seabass_fk_abun.csv")
  small_reef_abun   = read_csv("prep/data/rvc_trophic_groups/grouper_fk_abun.csv")
  
  grunt_etc_abun = rbind(grunt_abun,higher_reef_abun,small_reef_abun) #1999-2016 
  
  grunt_etc_abun = grunt_etc_abun %>%
    group_by(YEAR) %>%
    summarize(
      abundance_mean = mean(abundance),
      abundance_n = length(abundance),
      abundance_sd = sd(abundance),
      abundance_se = abundance_sd / sqrt(abundance_n),
      abundance_min = min(abundance),
      abundance_max = max(abundance))
  write_csv(grunt_etc_abun, "prep/data/rvc_trophic_groups/grunt_etc_abun.csv")
  
  #biomass data 
  grunt_bio = read_csv("prep/data/rvc_trophic_groups/grunt_fk_bio_merged.csv")
  higher_reef_bio = read_csv("prep/data/rvc_trophic_groups/higher_reef_fk_bio_merged.csv")
  small_reef_bio = read_csv("prep/data/rvc_trophic_groups/small_reef_fk_bio_merged.csv")
  
  grunt_etc_bio = rbind(grunt_bio,higher_reef_bio,opportunists_bio,small_reef_bio) #1999-2016 
  
  grunt_etc_bio = grunt_etc_bio %>%
    group_by(YEAR) %>%
    summarize(
      biomass_mean = mean(biomass),
      biomass_n = length(biomass),
      biomass_sd = sd(biomass),
      biomass_se = biomass_sd / sqrt(biomass_n),
      biomass_min = min(biomass),
      biomass_max = max(biomass))
  write_csv(grunt_etc_bio, "prep/data/rvc_trophic_groups/grunt_etc_bio.csv")
  
  #density data 
  grunt_den = read_csv("prep/data/rvc_trophic_groups/grunt_fk_den.csv")
  higher_reef_den = read_csv("prep/data/rvc_trophic_groups/higher_reef_fk_den.csv")
  small_reef_den = read_csv("prep/data/rvc_trophic_groups/small_reef_fk_den.csv")
  
  grunt_etc_den = rbind(grunt_den,higher_reef_den,small_reef_den) #1999-2016 
  
  grunt_etc_den = grunt_etc_den %>%
    group_by(YEAR) %>%
    summarize(
      density_mean = mean(density),
      density_n = length(density),
      density_sd = sd(density),
      density_se = density_sd / sqrt(density_n),
      density_min = min(density),
      density_max = max(density))
  write_csv(grunt_etc_den, "prep/data/rvc_trophic_groups/grunt_etc_den.csv")
  
  #diversity data 
  grunt_diversity = read_csv("prep/data/rvc_trophic_groups/grunt_fk_diversity.csv")
  higher_reef_diversity = read_csv("prep/data/rvc_trophic_groups/higher_reef_fk_diversity.csv")
  small_reef_diversity = read_csv("prep/data/rvc_trophic_groups/small_reef_fk_diversity.csv")
  
  grunt_etc_diversity = rbind(grunt_diversity,higher_reef_diversity,small_reef_diversity) #1994-2016 
  
  grunt_etc_diversity = grunt_etc_diversity %>%
    filter(protected_status != "0") %>%
    filter(protected_status != "1") %>%
    group_by(YEAR) %>%
    summarize(
      richness_mean = mean(richness),
      richness_n = length(richness),
      richness_sd = sd(richness),
      richness_se = richness_sd / sqrt(richness_n),
      richness_min = min(richness),
      richness_max = max(richness),
      simpson_mean = mean(simpson),
      simpson_n = length(simpson),
      simpson_sd = sd(simpson),
      simpson_se = simpson_sd / sqrt(simpson_n),
      simpson_min = min(simpson),
      simpson_max = max(simpson),
      shannon_mean = mean(shannon),
      shannon_n = length(shannon),
      shannon_sd = sd(shannon),
      shannon_se = shannon_sd / sqrt(shannon_n),
      shannon_min = min(shannon),
      shannon_max = max(shannon))
  
  write_csv(grunt_etc_diversity, "prep/data/rvc_trophic_groups/grunt_etc_diversity.csv")  
}

```

### Rays

| Scientific name               | Common name         | RVC code  |  
|:------------------------------|:--------------------|:----------|
|*Aetobatus narinari*	          | spotted eagle ray	  | AET NARI  |
|*Dasyatis americana*	          | southern stingray	  | DAS AMER  |
|*Echeneis naucrates*	          | sharksucker	        | ECH NAUC  |
|*Narcine bancroftii*	          | lesser electric ray	| NAR BANC  |
|*Rhinoptera bonasus*	          | cownosed ray	      | RHI BONA  |
|*Urobatis jamaicensis*	        | yellow stingray	    | URO JAMA  |

```{r rays}

if(!all(file.exists("prep/data/rvc_trophic_groups/rays_abun.csv",
                    "prep/data/rvc_trophic_groups/rays_den.csv",
                    "prep/data/rvc_trophic_groups/rays_bio.csv",
                    "prep/data/rvc_trophic_groups/rays_div.csv"))){
  
  rays_fk_abun_csv     = "prep/data/rays_fk_abun.csv"        #abundance
  rays_fk_den_csv      = "prep/data/rays_fk_den.csv"           #density
  rays_fk_bio_csv      = "prep/data/rays_fk_bio_merged.csv"    #biomass 
  rays_fk_diversity_csv= "prep/data/rays_fk_diversity.csv"   #diversity 
  
  if (!all(file.exists(
    rays_fk_abun_csv, rays_fk_den_csv, rays_fk_bio_csv,rays_fk_diversity_csv))){
    
    rays_fish_spp_list= c("HYP HARR", "ATH STIP", "JEN SPE.", "HAR JAGU", "ANC LYOL", "INE VITT", "HEM BRAS", "SAR AURI", "HAR HUME", "CHR ATHE")
    
    #abundance
    rays_fk_abun = getDomainAbundance(RVCdata_FK, rays_fish_spp_list)
    write_csv(rays_fk_abun, "prep/data/rvc_trophic_groups/rays_fk_abun.csv")
    
    #density
    rays_fk_den = getDomainDensity(RVCdata_FK, rays_fish_spp_list)
    write_csv(rays_fk_den, "prep/data/rvc_trophic_groups/rays_fk_den.csv")
    
    #biomass
    rays_fk_bio = getDomainBiomass(RVCdata_FK, rays_fish_spp_list, RVCdata_FK$taxonomic_data)
    write_csv(rays_fk_bio, "prep/data/rvc_trophic_groups/rays_fk_bio_merged.csv")
    
    #fk diversity metrics 
    rays_fk_abundance = rays_fk_abun %>% 
      select(YEAR, protected_status, SPECIES_CD, abundance) %>%  
      group_by(YEAR, protected_status) %>% #48 groups 
      nest(-YEAR) %>%  
      mutate(
        data_wide = map(data, ~ spread(data=.x, SPECIES_CD, abundance, fill = 0)), 
        #species richness 
        richness = map(
          data_wide, 
          function(x) specnumber(x)),
        #simpson diversity as effective number of species 
        simpson = map( 
          data_wide,
          function(x) 1/(1 - diversity(x, index = 'simpson'))),
        #shannon diversity as effective number of species 
        shannon = map(
          data_wide,
          function(x) exp(diversity(x, index = 'shannon')))) %>%
      unnest(richness, simpson, shannon)
    
    rays_fk_diversity = rays_fk_abundance %>%
      select(YEAR, protected_status, richness, shannon, simpson) 
    write_csv(rays_fk_diversity, "prep/data/rvc_trophic_groups/rays_fk_diversity.csv")
    
  } else{
    
    rays_fk_abun = read_csv("prep/data/rvc_trophic_groups/rays_fk_abun.csv")
    rays_abun = rays_fk_abun %>%
      group_by(YEAR) %>%
      summarize(
        abundance_mean = mean(abundance),
        abundance_n = length(abundance),
        abundance_sd = sd(abundance),
        abundance_se = abundance_sd / sqrt(abundance_n),
        abundance_min = min(abundance),
        abundance_max = max(abundance))
    write_csv(rays_abun, "prep/data/rvc_trophic_groups/rays_abun.csv")
    
    rays_fk_den = read_csv("prep/data/rvc_trophic_groups/rays_fk_den.csv")
    rays_den = rays_fk_den %>%
      group_by(YEAR) %>%
      summarize(
        density_mean = mean(density),
        density_n = length(density),
        density_sd = sd(density),
        density_se = density_sd / sqrt(density_n),
        density_min = min(density),
        density_max = max(density))
    write_csv(rays_den, "prep/data/rvc_trophic_groups/rays_den.csv")
    
    rays_fk_bio = read_csv("prep/data/rvc_trophic_groups/rays_fk_bio_merged.csv")
    rays_bio = rays_fk_bio %>%
      group_by(YEAR) %>%
      summarize(
        biomass_mean = mean(biomass),
        biomass_n = length(biomass),
        biomass_sd = sd(biomass),
        biomass_se = biomass_sd / sqrt(biomass_n),
        biomass_min = min(biomass),
        biomass_max = max(biomass))
    write_csv(rays_bio, "prep/data/rvc_trophic_groups/rays_bio.csv")
    
    rays_fk_diveristy = read_csv("prep/data/rvc_trophic_groups/rays_fk_diversity.csv") 
    rays_div = rays_fk_diveristy %>%
      filter(protected_status != "0") %>%
      filter(protected_status != "1") %>%
      select(YEAR, richness, shannon, simpson)
    write_csv(rays_div, "prep/data/rvc_trophic_groups/rays_div.csv")
  }
}
```

### Sharks

| Scientific name               | Common name         | RVC code  |  
|:------------------------------|:--------------------|:----------|
|*Carcharhinus falciformis*	    | silky shark	        | CAR FALC  |
|*Carcharhinus leucas*	        | bull shark	        | CAR LEUC  |
|*Carcharhinus limbatus*	      | blacktip shark	    | CAR LIMB  |
|*Carcharhinus obscurus*	      | dusky shark	        | CAR OBSC  |
|*Carcharhinus plumbeus*	      | sandbar shark	      | CAR PLUM  |
|*Carcharhinus perezii*	        | reef shark	        | CAR PERE  |
|*Galeocerdo cuvieri*	          | tiger shark	        | GAL CUVI  | 
|*Ginglymostoma cirratum*	      | nurse shark	        | GIN CIRR  |
|*Negaprion brevirostris*	      | lemon shark	        | Neg BREV  |
|*Sphyrna mokarran*	            | great hammerhead	  | SPH MOKA  |
|*Sphyrna tiburo*	              | bonnethead	        | SPH TIBU  |
|*Sphyrna lewini*	              | scalloped hammerhead| SPH LEWI  |

```{r sharks}

if(!all(file.exists("prep/data/rvc_trophic_groups/sharks_abun.csv",
                    "prep/data/rvc_trophic_groups/sharks_den.csv",
                    "prep/data/rvc_trophic_groups/sharks_bio_merged.csv",
                    "prep/data/rvc_trophic_groups/sharks_div.csv"))){
  
  sharks_fk_abun_csv     = "prep/data/sharks_fk_abun.csv"        #abundance
  sharks_fk_den_csv      = "prep/data/sharks_fk_den.csv"           #density
  sharks_fk_bio_csv      = "prep/data/sharks_fk_bio_merged.csv"    #biomass 
  sharks_fk_diversity_csv= "prep/data/sharks_fk_diversity.csv"   #diversity 
  
  if (!all(file.exists(
    sharks_fk_abun_csv, sharks_fk_den_csv, sharks_fk_bio_csv,sharks_fk_diversity_csv))){
    
    sharks_fish_spp_list= c("HYP HARR", "ATH STIP", "JEN SPE.", "HAR JAGU", "ANC LYOL", "INE VITT", "HEM BRAS", "SAR AURI", "HAR HUME", "CHR ATHE")
    
    #abundance
    sharks_fk_abun = getDomainAbundance(RVCdata_FK, sharks_fish_spp_list)
    write_csv(sharks_fk_abun, "prep/data/rvc_trophic_groups/sharks_fk_abun.csv")
    
    #density
    sharks_fk_den = getDomainDensity(RVCdata_FK, sharks_fish_spp_list)
    write_csv(sharks_fk_den, "prep/data/rvc_trophic_groups/sharks_fk_den.csv")
    
    #biomass
    sharks_fk_bio = getDomainBiomass(RVCdata_FK, sharks_fish_spp_list, RVCdata_FK$taxonomic_data)
    write_csv(sharks_fk_bio, "prep/data/rvc_trophic_groups/sharks_fk_bio_merged.csv")
    
    #fk diversity metrics 
    sharks_fk_abundance = sharks_fk_abun %>% 
      select(YEAR, protected_status, SPECIES_CD, abundance) %>%  
      group_by(YEAR, protected_status) %>% #48 groups 
      nest(-YEAR) %>%  
      mutate(
        data_wide = map(data, ~ spread(data=.x, SPECIES_CD, abundance, fill = 0)), 
        #species richness 
        richness = map(
          data_wide, 
          function(x) specnumber(x)),
        #simpson diversity as effective number of species 
        simpson = map( 
          data_wide,
          function(x) 1/(1 - diversity(x, index = 'simpson'))),
        #shannon diversity as effective number of species 
        shannon = map(
          data_wide,
          function(x) exp(diversity(x, index = 'shannon')))) %>%
      unnest(richness, simpson, shannon)
    
    sharks_fk_diversity = sharks_fk_abundance %>%
      select(YEAR, protected_status, richness, shannon, simpson) 
    write_csv(sharks_fk_diversity, "prep/data/rvc_trophic_groups/sharks_fk_diversity.csv")
    
  } else{
    
    sharks_fk_abun = read_csv("prep/data/rvc_trophic_groups/sharks_fk_abun.csv")
    sharks_abun = sharks_fk_abun %>%
      group_by(YEAR) %>%
      summarize(
        abundance_mean = mean(abundance),
        abundance_n = length(abundance),
        abundance_sd = sd(abundance),
        abundance_se = abundance_sd / sqrt(abundance_n),
        abundance_min = min(abundance),
        abundance_max = max(abundance))
    write_csv(sharks_abun, "prep/data/rvc_trophic_groups/sharks_abun.csv")
    
    sharks_fk_den = read_csv("prep/data/rvc_trophic_groups/sharks_fk_den.csv")
    sharks_den = sharks_fk_den %>%
      group_by(YEAR) %>%
      summarize(
        density_mean = mean(density),
        density_n = length(density),
        density_sd = sd(density),
        density_se = density_sd / sqrt(density_n),
        density_min = min(density),
        density_max = max(density))
    write_csv(sharks_den, "prep/data/rvc_trophic_groups/sharks_den.csv")
    
    sharks_fk_bio = read_csv("prep/data/rvc_trophic_groups/sharks_fk_bio_merged.csv")
    sharks_bio = sharks_fk_bio %>%
      group_by(YEAR) %>%
      summarize(
        biomass_mean = mean(biomass),
        biomass_n = length(biomass),
        biomass_sd = sd(biomass),
        biomass_se = biomass_sd / sqrt(biomass_n),
        biomass_min = min(biomass),
        biomass_max = max(biomass))
    write_csv(sharks_bio, "prep/data/rvc_trophic_groups/sharks_bio.csv")
    
    sharks_fk_diveristy = read_csv("prep/data/rvc_trophic_groups/sharks_fk_diversity.csv") 
    sharks_div = sharks_fk_diveristy %>%
      filter(protected_status != "0") %>%
      filter(protected_status != "1") %>%
      select(YEAR, richness, shannon, simpson)
    write_csv(sharks_div, "prep/data/rvc_trophic_groups/sharks_div.csv")
  }
}
```

### Lionfish

| Scientific name               | Common name         | RVC code  |  
|:------------------------------|:--------------------|:----------|
|*Pterois volitans*             | red lionfish        | PTE VOLI  |

```{r lionfish}

if(!all(file.exists("prep/data/rvc_trophic_groups/lionfish_abun.csv",
                    "prep/data/rvc_trophic_groups/lionfish_den.csv",
                    "prep/data/rvc_trophic_groups/lionfish_bio_merged.csv",
                    "prep/data/rvc_trophic_groups/lionfish_div.csv"))){
  
  lionfish_fk_abun_csv     = "prep/data/lionfish_fk_abun.csv"        #abundance
  lionfish_fk_den_csv      = "prep/data/lionfish_fk_den.csv"           #density
  lionfish_fk_bio_csv      = "prep/data/lionfish_fk_bio_merged.csv"    #biomass 
  lionfish_fk_diversity_csv= "prep/data/lionfish_fk_diversity.csv"   #diversity 
  
  if (!all(file.exists(
    lionfish_fk_abun_csv, lionfish_fk_den_csv, lionfish_fk_bio_csv,lionfish_fk_diversity_csv))){
    
    lionfish_fish_spp_list= c("HYP HARR", "ATH STIP", "JEN SPE.", "HAR JAGU", "ANC LYOL", "INE VITT", "HEM BRAS", "SAR AURI", "HAR HUME", "CHR ATHE")
    
    #abundance
    lionfish_fk_abun = getDomainAbundance(RVCdata_FK, lionfish_fish_spp_list)
    write_csv(lionfish_fk_abun, "prep/data/rvc_trophic_groups/lionfish_fk_abun.csv")
    
    #density
    lionfish_fk_den = getDomainDensity(RVCdata_FK, lionfish_fish_spp_list)
    write_csv(lionfish_fk_den, "prep/data/rvc_trophic_groups/lionfish_fk_den.csv")
    
    #biomass
    lionfish_fk_bio = getDomainBiomass(RVCdata_FK, lionfish_fish_spp_list, RVCdata_FK$taxonomic_data)
    write_csv(lionfish_fk_bio, "prep/data/rvc_trophic_groups/lionfish_fk_bio_merged.csv")
    
    #fk diversity metrics 
    lionfish_fk_abundance = lionfish_fk_abun %>% 
      select(YEAR, protected_status, SPECIES_CD, abundance) %>%  
      group_by(YEAR, protected_status) %>% #48 groups 
      nest(-YEAR) %>%  
      mutate(
        data_wide = map(data, ~ spread(data=.x, SPECIES_CD, abundance, fill = 0)), 
        #species richness 
        richness = map(
          data_wide, 
          function(x) specnumber(x)),
        #simpson diversity as effective number of species 
        simpson = map( 
          data_wide,
          function(x) 1/(1 - diversity(x, index = 'simpson'))),
        #shannon diversity as effective number of species 
        shannon = map(
          data_wide,
          function(x) exp(diversity(x, index = 'shannon')))) %>%
      unnest(richness, simpson, shannon)
    
    lionfish_fk_diversity = lionfish_fk_abundance %>%
      select(YEAR, protected_status, richness, shannon, simpson) 
    write_csv(lionfish_fk_diversity, "prep/data/rvc_trophic_groups/lionfish_fk_diversity.csv")
    
  } else{
    
    lionfish_fk_abun = read_csv("prep/data/rvc_trophic_groups/lionfish_fk_abun.csv")
    lionfish_abun = lionfish_fk_abun %>%
      group_by(YEAR) %>%
      summarize(
        abundance_mean = mean(abundance),
        abundance_n = length(abundance),
        abundance_sd = sd(abundance),
        abundance_se = abundance_sd / sqrt(abundance_n),
        abundance_min = min(abundance),
        abundance_max = max(abundance))
    write_csv(lionfish_abun, "prep/data/rvc_trophic_groups/lionfish_abun.csv")
    
    lionfish_fk_den = read_csv("prep/data/rvc_trophic_groups/lionfish_fk_den.csv")
    lionfish_den = lionfish_fk_den %>%
      group_by(YEAR) %>%
      summarize(
        density_mean = mean(density),
        density_n = length(density),
        density_sd = sd(density),
        density_se = density_sd / sqrt(density_n),
        density_min = min(density),
        density_max = max(density))
    write_csv(lionfish_den, "prep/data/rvc_trophic_groups/lionfish_den.csv")
    
    lionfish_fk_bio = read_csv("prep/data/rvc_trophic_groups/lionfish_fk_bio_merged.csv")
    lionfish_bio = lionfish_fk_bio %>%
      group_by(YEAR) %>%
      summarize(
        biomass_mean = mean(biomass),
        biomass_n = length(biomass),
        biomass_sd = sd(biomass),
        biomass_se = biomass_sd / sqrt(biomass_n),
        biomass_min = min(biomass),
        biomass_max = max(biomass))
    write_csv(lionfish_bio, "prep/data/rvc_trophic_groups/lionfish_bio.csv")
  }
}

```

- trophic pyramid 

```{r trophic pyramid}

if(!file.exists("prep/data/rvc_trophic_level/trophic_level_biomass.csv")){ 
  
  trophic_levels = read_csv('prep/data/rvc_trophic_level/RVC_trophic_level.csv')
  
  trophic_level = trophic_levels %>%
    select(RVC_CODE, Trophic_Level) %>%  #391 species 
    mutate(
      TL = # 3 trophic levels 
        ifelse(Trophic_Level == 2, "TL2",
               ifelse(Trophic_Level == 3, "TL3",
                      ifelse(Trophic_Level == 4, "TL4", NA))),
      SPECIES_CD = RVC_CODE) %>% 
    arrange(SPECIES_CD) %>% 
    select(SPECIES_CD, TL) 
  
  trophic_level_bio = getDomainBiomass(RVCdata_FK, trophic_level$RVC_CODE, RVCdata_FK$taxonomic_data)
  write_csv(trophic_level_bio, "prep/data/rvc_trophic_level/trophic_level_bio.csv")
  
  trophic_level_bio = trophic_level_bio %>%
  select(YEAR, biomass, SPECIES_CD) %>% 
  left_join(trophic_level, trophic_level_bio, by = 'SPECIES_CD') %>% 
  #trophic_level_biomass = trophic_level_bio %>% 
  group_by(YEAR, TL) %>%
  summarise(
  ntot = sum(biomass)) %>% 
  spread(TL, ntot)
  
  write_csv(trophic_level_bio, "prep/data/rvc_trophic_level/trophic_level_biomass")
  
}

```

- exploited reef fish 

```{r exploited reef fish}

if(!all(file.exists("prep/data/rvc_exploited_species/exploited_den.csv",
                    "prep/data/rvc_exploited_species/exploited_bio.csv",
                    "prep/data/rvc_exploited_species/exploited_abun.csv"))){
  
  exploited_fk_abun_csv = "prep/data/rvc_exploited_species/exploited_fk_abun.csv"
  exploited_fk_den_csv  = "prep/data/rvc_exploited_species/exploited_fk_den.csv"
  exploited_fk_bio_csv  = "prep/data/rvc_exploited_species/exploited_fk_bio.csv"
  
}else{ 
  if(!all(file.exists(exploited_fk_abun_csv, exploited_fk_den_csv, exploited_fk_bio_csv))){
    
    exploited_spp_list= c(
      "CEP CRUE", #Graysby
      "EPI ITAJ", #Goliath Grouper 
      "EPI MORI", #Red Grouper
      "EPI STRI", #Nassau Grouper 
      "LAC MAXI", #Hogfish
      "LUT ANAL", #Mutton Snapper
      "LUT GRIS", #Grey Snapper    
      "MYC BONA", #Black Grouper
      "OCY CHRY" #Yellowtail Snapper 
    )
    #abundance
    exploited_fk_abun <- getDomainAbundance(RVCdata_FK, species = exploited_spp_list, merge_protected = F)
    write_csv(exploited_fk_abun,"prep/data/rvc_exploited_species/exploited_fk_abun.csv")
    
    #density 
    exploited_fk_den <- getDomainDensity(RVCdata_FK, species = exploited_spp_list, merge_protected = F)
    write_csv(exploited_fk_den,"prep/data/rvc_exploited_species/exploited_fk_den.csv")
    
    #biomass 
    exploited_fk_bio <- getDomainBiomass(RVCdata_FK, species = exploited_spp_list, growth_parameters = RVCdata_FK$taxonomic_data, merge_protected = F)
    write_csv(exploited_fk_bio,"prep/data/rvc_exploited_species/exploited_fk_bio.csv")
  }
  
  exploited_den = read_csv("prep/data/rvc_exploited_species/exploited_fk_den.csv")
  
  exploited_density = exploited_den %>%
    mutate(
      protected_status = recode(
        protected_status, '0'='unprotected', '1'='protected'), 
      common_name = recode(
        SPECIES_CD,
        'CEP CRUE' = 'Graysby',
        'EPI ITAJ' = 'Goliath_Grouper',
        'EPI MORI' = 'Red_Grouper',
        'EPI STRI' = 'Nassau_Grouper',
        'LAC MAXI' = 'Hogfish',
        'LUT ANAL' = 'Mutton_Snapper',
        'LUT GRIS' = 'Grey_Snapper',
        'MYC BONA' = 'Black_Grouper',
        'OCY CHRY' = 'Yellowtail_Snapper')) %>% 
    select(YEAR, common_name, protected_status, density) %>%
    group_by(YEAR, common_name, protected_status) %>%
    filter(protected_status != "all") 
  
  write_csv(exploited_density, "prep/data/rvc_exploited_species/exploited_den.csv")
  
  exploited_bio = read_csv("prep/data/rvc_exploited_species/exploited_fk_bio.csv")
  
  exploited_biomass = exploited_bio %>%
    mutate(
      protected_status = recode(
        protected_status, '0'='unprotected', '1'='protected'), 
      common_name = recode(
        SPECIES_CD,
        'CEP CRUE' = 'Graysby',
        'EPI ITAJ' = 'Goliath_Grouper',
        'EPI MORI' = 'Red_Grouper',
        'EPI STRI' = 'Nassau_Grouper',
        'LAC MAXI' = 'Hogfish',
        'LUT ANAL' = 'Mutton_Snapper',
        'LUT GRIS' = 'Grey_Snapper',
        'MYC BONA' = 'Black_Grouper',
        'OCY CHRY' = 'Yellowtail_Snapper')) %>% 
    select(YEAR, common_name, protected_status, biomass) %>%
    group_by(YEAR, common_name, protected_status) %>%
    filter(protected_status != "all") 
  
  write_csv(exploited_biomass, "prep/data/rvc_exploited_species/exploited_bio.csv")
  
  exploited_abun = read_csv("prep/data/rvc_exploited_species/exploited_fk_abun.csv")
  
  exploited_abundance = exploited_abun %>%
    mutate(
      protected_status = recode(
        protected_status, '0'='unprotected', '1'='protected'), 
      common_name = recode(
        SPECIES_CD,
        'CEP CRUE' = 'Graysby',
        'EPI ITAJ' = 'Goliath_Grouper',
        'EPI MORI' = 'Red_Grouper',
        'EPI STRI' = 'Nassau_Grouper',
        'LAC MAXI' = 'Hogfish',
        'LUT ANAL' = 'Mutton_Snapper',
        'LUT GRIS' = 'Grey_Snapper',
        'MYC BONA' = 'Black_Grouper',
        'OCY CHRY' = 'Yellowtail_Snapper')) %>% 
    select(YEAR, common_name, protected_status, abundance) %>%
    group_by(YEAR, common_name, protected_status) %>%
    filter(protected_status != "all") 
  
  write_csv(exploited_abundance, "prep/data/rvc_exploited_species/exploited_abun.csv")
  
}

```

- reef fish biodiversity 

```{r reef fish biodiversity}

if(!file.exists("prep/data/domain_div.csv")){

  domain_abun = read_csv("C:/Users/mhepn/Documents/RVC/big_csv/abundance_domain/domain_fk_abun.csv") #1307 KB

  domain_div = domain_abun %>%
    filter(protected_status == "all") %>%
    select(YEAR, SPECIES_CD, abundance) %>%
    group_by(YEAR) %>%
    nest(-YEAR) %>%
    mutate(
      data_wide = map(data, ~ spread(data=.x, SPECIES_CD, abundance, fill =0)),
      richness = map(data_wide, function(x) specnumber(x)),
      simpson  = map(data_wide, function(x) 1/(1 - diversity(x, index = 'simpson'))),
      shannon = map(data_wide,  function(x) exp(diversity(x,  index = 'shannon')))) %>% 
    unnest(richness, simpson, shannon)

  domain_div = domain_div %>%
    select(YEAR, richness, simpson, shannon)

  write_csv(domain_div, "prep/data/domain_div.csv")
}

```

- queen conch

```{r queen conch}

queen_conch = read_csv("prep/data/queen_conch/QueenConchData.csv")

```

- caribbean spiny lobster 

```{r caribbean spiny lobster}

```

- stony coral
- octocoral
- macroalgae 
- sponges 

```{foundation species}

```