---
title: "Prep Infographic Data" 
author: "Ben Best"
date: "`r Sys.Date()`"
output:
  html_document:
    df_print: paged
    toc: true
    toc_float: true
    code_folding: hide
---

## Load Software

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r load infographix, message=F, warning=F}
if (!require('devtools')) install.packages('devtools')
if (!require('infographiq')) devtools::install_github('marinebon/infographiq')
#library(infographiq)
# library(devtools); load_all('../../infographiq')
#session_info()

library(tidyverse)
library(xml2)
library(stringr)

if ( basename(getwd()) != 'prep' ) setwd('prep')

dir_svg = '../svg'
indicators_csv = '../plot_indicators.csv'
```

## Read svg & csv

### plot_indicators.csv

```{r plot_indicators, message=F}
# read_csv('https://raw.githubusercontent.com/mhepner90/RVC/master/SDG14/infographic_csv_indicators.csv') %>%
#    write_csv(indicators_csv)
ind_d = read_csv(indicators_csv)

ind_d %>%
  group_by(svg_id) %>%
  summarize(
    n = n())
```

### *.svg

```{r svg_elements}
svg_ids = function(svg){
  ids = read_xml(svg) %>%
    xml_find_all("//svg:g") %>%
    xml_attr('id')
  
  ids = ids[!str_detect(ids, '^g\\d+$')]
  if (length(ids)==0){
    return(NA)
  } else {
    return(ids)
  }
}

svg_tbl = tibble()
for (svg in list.files(dir_svg, '.*\\.svg$')){
  svg_tbl = svg_tbl %>%
    bind_rows(
      data_frame(
        svg   = svg,
        svg_id = svg_ids(
          file.path(dir_svg, svg))))
}

svg_tbl %>%
  arrange(svg, svg_id)
```

## Check svg_id matches

### *.svg missing plot_indicators.csv

```{r svg_elements missing plot_indicators}
svg_tbl %>%
  anti_join(ind_d, by='svg_id') %>%
  arrange(svg, svg_id)
```

### *.svg missing plot_indicators.csv csv_url

```{r svg_elements missing plot_indicator.url_csv}
svg_tbl %>%
  inner_join(ind_d, by='svg_id') %>%
  filter(is.na(csv_url)) %>%
  select(svg, svg_id, title, region, csv_url) %>%
  arrange(svg, svg_id, title, region)
```

### plot_indicators.csv missing *.svg

```{r plot_indicators missing svg_elements}
ind_d %>%
  anti_join(svg_tbl, by='svg_id') %>%
  select(svg_id, title, region, csv_url) %>%
  arrange(svg_id, title, region, csv_url) # %>% View()
```

## Query taxa groupings

### snapper / grouper ?

Here's any svg_id with the word snapper or grouper in it. Isn't this often thought of as the "snapper grouper complex", so should we aggregate these time series? 

```{r}
ind_d %>%
  filter(str_detect(svg_id, regex('(snapper|grouper)', ignore_case=T))) %>%
  group_by(svg_id) %>%
  summarize(
    n_rows = n())
```


## Create Site

```{r, eval=F}
load_all('../../infographiq') # reload package from source

create_site()
```



...

